---
title: "Kansas Secretary of State Monthly Inactive Voter Plots by County"
author: "Earl F Glynn<br><small>watchdoglab.substack.com/</small>"
date: "<small>`r Sys.Date()`</small>"
output:
  html_document:
    code_download: true
    theme: cerulean
    toc: yes
    toc_depth:  3
    toc_float:
      collapsed:  yes
      smooth_scroll: yes
    number_sections: yes
    code_folding:  show
---

Kansas Secretary of State "Inactive" Data: [Election Statistics Data, Voter Registration, Monthly County  Totals](https://sos.ks.gov/elections/election-statistics-data.html)

See data problems that should be addressed by Kansas Secretary of State in `Inactive-Stats.Rmd`.

```{r setup, echo = FALSE}
# http://biostat.mc.vanderbilt.edu/wiki/Main/KnitrHtmlTemplate
require(Hmisc)    # provides knitrSet and other functions
knitrSet(lang = 'markdown',   # If using blogdown: knitrSet(lang='blogdown')
         fig.align = 'left',
         w = 6.5,
         h = 4.5,
         cache = FALSE)
```

`r hidingTOC(buttonLabel = "Outline")`

```{r startYourEngines, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  comment = NA)

time.1 <- Sys.time()
```

# Setup {.tabset .tabset-fade .tabset-pills}

## {.active}

## Constants

```{r}
plotCaptionLeft  <- "Kansas Secretary of State: Online Election Statistics (circles) [many months missing]; Voter files (triangles)"
plotCaptionRight <- paste("watchdoglab.substack.com", format(Sys.time(), "%Y-%m-%d"))
```       

## Packages

```{r Packages}
library(tidyverse)
library(lubridate)
```

Display formatting

```{r Display}
library(kableExtra)  # kable_styling
```

I/O

```{r IO}
library(readxl)      # read_xlsx
library(readr)       # write_csv
library(writexl)     # write_xlsx
```

## ggplot theme

[ggplot Themes Gallery](https://www.datanovia.com/en/blog/ggplot-themes-gallery/)

```{r ggplotTheme}
theme_set(
          theme_minimal() +

          theme(
                axis.text             = element_text(size = 10),
                axis.title            = element_text(size = 14),

                legend.position       = "bottom",

                plot.caption          = element_text(hjust = c(0.0,1.0),
                                                     size = 10),
                plot.caption.position = "plot",

                plot.title.position   = "plot",

                strip.background      = element_rect(fill = "aliceblue"),
                strip.text            = element_text(size = 14),

                plot.title            = element_text(size = 14, color = "blue")
               )
         )
```
                                                  
## Helper functions

```{r Helpers}
Show <- function(data, caption="", bigMark="",
                 height = NULL, width = NULL, ...)
{
  data                                       |>
  kable("html", caption=caption,
        format.args=list(big.mark=bigMark))  |>
  kable_styling(bootstrap_options=c("striped", "bordered", "condensed"),
                position="left",
                full_width=FALSE, ...)       |>
  scroll_box(height = height, width = width)
}
```

# Voter file stats

```{r}
voterSummary <- 
  read_xlsx("Kansas-Voter-Counts-2023-02-13-thru-2025-01-02.xlsx") |>
  select(County, Date, InactivePercent) |>
  mutate(Date = ymd(Date))
```

# List of summary files

"Long" format, suitable for ggplot

```{r}
filename <- "Kansas-SoS-Inactive-Voter-Stats-2021-06-01-thru-2025-02-01.xlsx"

countySummary <- 
  read_xlsx(filename) |>
  mutate(Date = ymd(Date))

glimpse(countySummary)
```

```{r}
fivenum(countySummary$InactivePercent)
```

```{r}
boxplot(countySummary$InactivePercent)
```


```{r}
countySummary |>
  filter(InactivePercent > 20) |>
  Show()
```


# All Counties

```{r CountyLines, fig.width = 10, fig.height = 6}
stateStats   <- countySummary |> filter(County == "Kansas")
stateVoters  <- voterSummary  |> filter(County == "Kansas")
stateColor   <- "royalblue"

allCountiesPlot <- 
  ggplot() +
    geom_line(data = countySummary, 
              aes(x = Date,
                  y = InactivePercent,
                  group = County),
              color = "grey")                                        +
    
    geom_line(data = stateStats,
              aes(x = Date,
                  y = InactivePercent),
              color = stateColor,
              linewidth = 1)                                         +
    
     geom_point(data = stateStats,
              aes(x = Date,
                  y = InactivePercent),
              color = stateColor,
              size = 3)                                              +
    
    geom_point(data = voterSummary |> filter(County == "Kansas"),
              aes(x = Date,
                  y = InactivePercent),
              color = stateColor, fill= stateColor,
              size = 3,
              shape = 25)                                             +
    
    
    scale_y_continuous(limits = c(0, 23),
                       breaks = c(0, 5, 10, 15, 20),
                       labels = c("0%", "5%", "10%", "15%", "20%"))  +
    
    scale_x_date(limits = c(ymd("2021-07-01"), ymd("2025-03-01")),
                 date_breaks       = "3 months",
                 date_minor_breaks = "1 month",
                 date_labels = "%Y\n%b",
                 expand = c(0,0))                                    +
    
    labs(title = "Percent Inactive Voters in Kansas",
         subtitle = "Grey lines show inactive voter rate profiles for each of the 105 Kansas counties",
         x = "Date",
         y = "Percent Inactive Voters",
         caption = c(plotCaptionLeft, plotCaptionRight))  +
     theme(plot.title         = element_text(size = 14, color = stateColor),
           plot.subtitle      = element_text(size = 14, color = "grey"),
           axis.ticks.x       = element_line(),
           panel.grid.major.x = element_line(linewidth = 0.5, color = "grey"),
           panel.grid.minor.y = element_blank(),
           panel.grid.major.y = element_line(linewidth = 0.5, color = "grey"))

print(allCountiesPlot)
```

Save as PDF and SVG files

```{r}
pdf("Kansas-Inactive-Voter-Rates-All-Counties.pdf", 
    width = 10, height = 6)  
print(allCountiesPlot)
dev.off()
```

```{r}
svg("Kansas-Inactive-Voter-Rates-All-Counties.svg", 
    width = 10, height = 6, pointsize = 8)  
print(allCountiesPlot)
dev.off()
```


# Single County

```{r targetCountyFunction}
PlotCounty <- function(targetCounty)
{
  targetStats  <- countySummary |> filter(County == targetCounty)
  targetVoters <- voterSummary  |> filter(County == targetCounty)
  targetColor  <- "firebrick"
  
  countyPlot <- 
    ggplot() +
     # geom_line(data = targetStats, 
     #            aes(x = Date,
     #                y = InactivePercent,
     #                group = County),
     #            color = targetColor)                                 +
      
      geom_point(data = targetStats,
                aes(x = Date,
                    y = InactivePercent),
                color = targetColor,
                size = 3)                                              +
      
      geom_point(data = targetVoters,
                aes(x = Date,
                    y = InactivePercent),
                color = targetColor, fill= targetColor,
                size = 3,
                shape = 25)                                            +
      
     
     # geom_line(data = stateStats,
     #            aes(x = Date,
     #                y = InactivePercent),
     #            color = "blue",
     #            linewidth = 1)                                       +
      
       geom_point(data = stateStats,
                aes(x = Date,
                    y = InactivePercent),
                color = stateColor,
                size = 3)                                              +
      
      geom_point(data = stateVoters,
                aes(x = Date,
                    y = InactivePercent),
                color = stateColor, fill= stateColor,
                size = 3,
                shape = 25)                                             +
      
      
      scale_y_continuous(limits = c(0, 23),
                         breaks = c(0, 5, 10, 15, 20),
                         labels = c("0%", "5%", "10%", "15%", "20%"))  +
      
      scale_x_date(limits = c(ymd("2021-07-01"), ymd("2025-03-01")),
                   date_breaks       = "3 months",
                   date_minor_breaks = "1 month",
                   date_labels = "%Y\n%b",
                   expand = c(0,0))                                    +
      
      labs(title    = "Percent Inactive Voters in Kansas",
           subtitle = paste(targetCounty, "County"),
           x = "Date",
           y = "Percent Inactive Voters",
           caption = c(plotCaptionLeft, plotCaptionRight))  +
           theme(plot.subtitle      = element_text(size = 14, color = targetColor),
                 plot.title         = element_text(size = 14, color = stateColor),
                 axis.ticks.x       = element_line(),
                 panel.grid.major.x = element_line(linewidth = 0.5, color = "grey"),
                 panel.grid.minor.y = element_blank(),
                 panel.grid.major.y = element_line(linewidth = 0.5, color = "grey"))

  return(countyPlot)
}
```

## Example County

```{r SingleCounty, fig.width = 10, fig.height = 6}
countyPlot <- PlotCounty("Sumner")
print(countyPlot)
```
```{r}
counties <- unique(countySummary$County) |> head(105)   # Remove final "Kansas" rwo
counties
```

## All in single PDF

```{r}
pdf("Kansas-Inactive-Voter-Rates-By-County.pdf", 
    width = 10, height = 6)  
  
  for (county in counties)
  {
    countyPlot <- PlotCounty(county)
    print(countyPlot)
  }

  print(allCountiesPlot)
  
dev.off()
```

## Separate SVG files in "County" folder

```{r}
DIR <- "County"
```

```{r}
if (!file.exists(DIR) )  
{
  dir.create(DIR)
}
```

```{r}
for (county in counties)
{
  countyPlot <- PlotCounty(county)
    
  svg(paste0(DIR, "/", county, "-County-Inactive-Voter-Rates.svg"), 
      width = 10, height = 6, pointsize = 8) 
    print(countyPlot)
  dev.off()
}
```

# Epilog {.tabset .tabset-fade .tabset-pills}

## {.active}

## Session Info

```{r devtoolsSessionInfo}
devtools::session_info()
```

</div>

```{r epilogDeltaTime, echo=FALSE}
time.2 <- Sys.time()
processingTime <- paste("Processing time:", sprintf("%.1f",
                        as.numeric(difftime(time.2,
                                            time.1, units="secs"))), "secs\n")
```

`r processingTime`
`r format(time.2, "%Y-%m-%d %H%M")`

