---
title: "Kansas Secretary of State Monthly Inactive Voter Stats"
author: "Earl F Glynn<br><small>watchdoglab.substack.com/</small>"
date: "<small>`r Sys.Date()`</small>"
output:
  html_document:
    code_download: true
    theme: cerulean
    toc: yes
    toc_depth:  3
    toc_float:
      collapsed:  yes
      smooth_scroll: yes
    number_sections: yes
    code_folding:  show
---

Kansas Secretary of State "Inactive" Data: [Election Statistics Data, Voter Registration, Monthly County  Totals](https://sos.ks.gov/elections/election-statistics-data.html)

Secretary of State should address:

* consistency in filenames across years 
  (new filename format in 2025:  "2025.01.01 Percentage of Inactive Voters by County.xlsx)
  
* consistency in column headers across files 
  ("COUNTY" was "14,464" in 2021-08-01,
                "16,022" in 2021-09-01,
                "16,000" in 2021-11-01, 2021-12-01,
                "16,027" in 2022-01-01,
                missing  in 2022-08-01 and later)
                
* missing "inactive" files: 

**2021**

  2021-10
  
**2022**

  2022-02
  2022-03
  2022-04
  2022-05
  2022-06
  2022-07
  2022-09
  2022-10
  2022-11
  2022-12

**2023**

  2023-01
  2023-02
  2023-03
  2023-04
  2023-05
  2023-06
  2023-07
  2023-08
  2023-09
  2023-10

**2024**

  2024-05
  2024-07
  2024-08
  2024-09
  2024-10
  2024-11
  2024-12
  
**2025**

  2025-03
  

```{r setup, echo = FALSE}
# http://biostat.mc.vanderbilt.edu/wiki/Main/KnitrHtmlTemplate
require(Hmisc)    # provides knitrSet and other functions
knitrSet(lang = 'markdown',   # If using blogdown: knitrSet(lang='blogdown')
         fig.align = 'left',
         w = 6.5,
         h = 4.5,
         cache = FALSE)
```

`r hidingTOC(buttonLabel = "Outline")`

```{r startYourEngines, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  comment = NA)

time.1 <- Sys.time()
```

# Setup {.tabset .tabset-fade .tabset-pills}

## {.active}

## Packages

```{r Packages}
library(tidyverse)
library(lubridate)
```

Display formatting

```{r Display}
library(kableExtra)  # kable_styling
```

I/O

```{r IO}
library(readxl)      # read_xlsx
library(readr)       # write_csv
library(writexl)     # write_xlsx
```

## Helper functions

```{r Helpers}
Show <- function(data, caption="", bigMark="",
                 height = NULL, width = NULL, ...)
{
  data                                       |>
  kable("html", caption=caption,
        format.args=list(big.mark=bigMark))  |>
  kable_styling(bootstrap_options=c("striped", "bordered", "condensed"),
                position="left",
                full_width=FALSE, ...)       |>
  scroll_box(height = height, width = width)
}
```

# List of summary files

```{r Names}
options(width = 50)
summaryFile <- list.files(path       = ".",
                          pattern    = "Percentage-of-Inactive|Percentage of Inactive",
                          all.files  = FALSE,   # don't want "." and ".."
                          full.names = FALSE,   # don't want path
                          recursive  = FALSE,
                          include.dirs = FALSE)  
summaryFile
```

Note in 2025 filenames "." used instead of "-" and " " used instead of "-"

Fix dates for consistency in any labels

```{r}
fileDate <- 
  str_sub(summaryFile, 1, 10) |>
  str_replace_all("\\.", "-")

fileDate
```

# Function to read a single inactive file

Extra rows need to be removed. 

```{r readFile}
readFile <- function(index)
{
  raw <- 
    read_xlsx(summaryFile[index]) |>
    select(-4)   # drop percentage; will compute below with additional digit
  
  # First column is inconsistently named and absent in later files.
  names(raw) <- c("County", "Voters", "Inactive")
 
  # Compute InactivePercent
  raw <- 
    raw                                                     |> 
    mutate(
            Date   = fileDate[index],
            Active = Voters - Inactive,
            InactivePercent = round(100 * Inactive / Voters, 2)
           )                                                |>
    relocate(Date,            .after = "County")            |>
    relocate(Active,          .after = "Voters")            |>
    relocate(InactivePercent, .after = "Inactive")       
  
  raw$County[nrow(raw)] <- "Kansas"    # Missing label for statewide values
  
  raw <- raw[-c(1, 107), ]   # Delete unwanted blank rows
  
  return(raw)
}
```

# Read all inactive files

```{r}
countySummary <- 
  1:length(summaryFile) |>
  purrr::map_dfr(readFile)
 
glimpse(countySummary)
```

Statewide values

```{r}
countySummary |> filter(County == "Kansas")  |> Show(bigMark = ",")
```

```{r}
write_xlsx(countySummary, 
           paste0("Kansas-SoS-Inactive-Voter-Stats-", 
                  head(fileDate,1),
                  "-thru-",
                  tail(fileDate,1),
                  ".xlsx"))
```

```{r}
table(countySummary$Date)
```

# Epilog {.tabset .tabset-fade .tabset-pills}

## {.active}

## Session Info

```{r devtoolsSessionInfo}
devtools::session_info()
```

</div>

```{r epilogDeltaTime, echo=FALSE}
time.2 <- Sys.time()
processingTime <- paste("Processing time:", sprintf("%.1f",
                        as.numeric(difftime(time.2,
                                            time.1, units="secs"))), "secs\n")
```

`r processingTime`
`r format(time.2, "%Y-%m-%d %H%M")`

